Embedded Meta-Learning: Toward more flexible deep-learning models

### 2019-08-11 writing/embedded_meta_learning_main.tex
Refactoring, clarifying

\paragraph{Meta-mapping:} However, the system is not limited to simply mapping inputs to outputs. From the perspective of our architecture, learning a meta-mapping between tasks is exactly analogous to learning a basic task. Anything that is embedded in $Z$ can be transformed using the same system. Because tasks are embedded in $Z$, this allows for meta-mappings using exactly the same $\mathcal{M}$ and $\mathcal{H}$. For example, suppose we have an embedding $z_{game1} \in Z$ for the task of playing some game. We can generate a meta-mapping embedding $z_{\text{meta}} \in Z$ from embeddings generated by the system when it is trying to win and lose various games: $z_{\text{meta}} = \mathcal{M}\left( \left\{\left((z_{game2},z_{game2,lose}\right), ... \right\}\right)$. We can generate a new task embedding $\hat{z}_{new} \in Z$:  

@@

Alternatively, we could map from language to a meta-mapping embedding, rather than inducing it from examples of the meta-mapping. This corresponds to the human ability to change behavior in response to instructions. The key feature of our architecture -- the fact that tasks, data, and language are all embedded in a shared space -- allows for substantial flexibility within a unified framework.

@@

\begin{subfigure}{0.5\textwidth}
\includegraphics[width=\textwidth]{figures/poly/meta_results.png}
\caption{From examples of mapping.}
\label{poly_meta_map_results_examples}
\end{subfigure}%
\begin{subfigure}{0.5\textwidth}
\includegraphics[width=\textwidth]{figures/poly/language_meta_results.png}
\caption{From language input.}
\label{poly_meta_map_results_language}
\caption{The system generalizes to apply learned meta-mappings to new polynomials, and to apply unlearned meta-mappings, either from examples (\subref{poly_meta_map_results_examples}) or from language cues (\subref{poly_meta_map_results_language}). The plots show the loss produced when evaluating the mapped embedding on the target task. For example, if the initial polynomial is $p(x) = x + 1$, and the meta-task is ``square,'' the loss would be evaluated by applying this meta-mapping to the embedding of $p(x)$ and evaluating how well the function specified by the mapped embedding regresses $p(x)^2 = x^2 + 2x + 1$. The solid line indicates optimal performance; the dashed line is chance (untrained). Error-bars are bootstrap 95\%-CIs.} 
\label{poly_meta_map_results}
%% Cut example if necessary
As a proof of concept, we first evaluated the system on the task of learning polynomials of degree $\leq 2$ in 4 variables (i.e. the task was to regress functions of the form $p: \mathbb{R}^4 \rightarrow \mathbb{R}$ where $p \in \mathcal{P}_2 \left(\mathbb{R}\right)$, though the model was given no prior inductive bias toward polynomial forms). For example, if $p(w,x,y,z) = x$, the model might see examples like $(-1,1,1,1; 1)$ and $(0.7, 2.1, 1.3, -4; 2.1)$, and be evaluated on its output for points like $(-1, -1.3, 0.5, 0.3)$. This yields an infinite family of base-level tasks (the vector space of all such polynomials), as well as many families of meta-mappings over tasks (for example, multiplying polynomials by a constant, squaring them, or permuting their input variables). This allows us to not only examine the ability of the system to learn to learn polynomials from data, but also to adapt its learned representations in accordance with these meta-tasks. Details of the architecture and training can be found in appendix \ref{app_detailed_methods}.\par
\paragraph{Basic meta-learning:} First, we show that the system is able to achieve the basic goal of learning a held-out polynomial from a few data points in Fig. \ref{poly_basic_results} (with good sample-efficiency, see supp. fig. \ref{supp_poly_sweep_results}). \par 
\paragraph{Meta-mapping (task $\rightarrow$ task):} Furthermore, the system is able to perform meta-mappings over polynomials in order to flexibly reconfigure its behavior, as shown in fig. \ref{poly_meta_map_results}. We train the system to perform a variety of mappings, for example switch the first two inputs of the polynomial, add 3 to the polynomial, or square the polynomial. We then test its ability to generalize to held-out mappings, for example a permutation of the polynomial inputs which it has not encountered before, or an additive shift in the polynomial which it has not been trained on. The system is both able to apply its learned meta-mappings to held-out tasks, and to apply meta-mappings it has not been trained on, simply by seeing examples of the mapping or receiving a natural-language cue to the meta-mapping. It performs slightly better from examples than language, since examples provide a richer description of the meta-mapping. \par 
A very simple solution to these problems is naturally suggested by our architecture. Specifically, task embeddings can be cached so that they don't have to be regenerated at each inference step. This also allows optimization of these embeddings without altering the other parameters in the architecture, thus allowing fine-tuning on a task without seeing more examples, and without interfering with performance on any other task \citep[cf.][]{Rumelhart1993, Lampinen2018a}. That is, we can see the meta-learning step as a ``warm start'' for an optimization procedure over embeddings that are cached in memory \citep[cf.][]{Kumaran2016}.\footnote{Note that the weights generated by the hyper network could be cached instead of the embeddings, which would be much more memory intensive but would allow even better tuning and would not require passing through the hyper network at inference time.} \par 
To test this idea, we pre-trained the system on 100 polynomial tasks, and then introduced 100 new tasks. We trained on these new tasks by starting from the meta-network's ``guess'' at the correct task embedding, and then optimizing this embedding without altering the other parameters. The results are shown in fig. \ref{poly_continual_results}. The meta-network embeddings offer good immediate performance, and substantially accelerate the optimization process, compared to a randomly-initialized embedding (see supp. fig. \ref{supp_poly_continual_results} for a more direct comparison). Furthermore, this ability to learn is due to training, not simply the expressiveness of the architecture, as is shown by attempting the same with an untrained network. \par

@@

@@

\begin{subfigure}{0.5\textwidth}
\includegraphics[width=\textwidth]{figures/meta_mapping.png}
\caption{From examples of mapping.}
\label{cards_meta_map_results_examples}
\end{subfigure}%
\begin{subfigure}{0.5\textwidth}
\includegraphics[width=\textwidth]{figures/language_meta_mapping.png}
\caption{From language input.}
\label{cards_meta_map_results_language}
\caption{The system generalizes to meta-mapping new tasks in the cards domain. The system is trained to do the meta-mappings shown here on a subset of its basic tasks, and is able to generalize these mappings to novel tasks it has not been trained on. For example, for the ``losers'' mapping, the sytem is trained to map games to their losers variants. When given a novel game, it is able to apply this mapping to guess how to play the losers variation. This plot shows the reward produced by taking the mapped embedding and playing the targeted game. The gray dashed line indicates random performance, while the colored dashed lines indicate performance if the system did not alter its behavior in response to the meta-mapping. The system generally exceeds these baselines, although the switch-suits baseline is more difficult with the targeted holdout. The meta-mappings can be cued either from examples (\subref{cards_meta_map_results_examples}) or from language cues (\subref{cards_meta_map_results_language}). Error-bars are bootstrap 95\%-CIs.}
\label{cards_meta_map_results}

@@

>>>>

\paragraph{Meta-mapping:} However, the system is not limited to simply mapping inputs to outputs. The fundamental insight of our paper is to show how basic tasks and meta-mappings can be treated homogenously, by allowing the network to recurse on its own representations. From the perspective of our architecture, learning a meta-mapping between tasks is exactly analogous to learning a basic task. Anything that is embedded in $Z$ can be transformed using the same system. Because tasks are embedded in $Z$ for basic meta-learning, this allows for meta-mappings using exactly the same $\mathcal{M}$ and $\mathcal{H}$ that we use for basic tasks. For example, suppose we have an embedding $z_{game1} \in Z$ for the task of playing some game. We can generate a meta-mapping embedding $z_{\text{meta}} \in Z$ from embeddings generated by the system when it is trying to win and lose various games: $z_{\text{meta}} = \mathcal{M}\left( \left\{\left((z_{game2},z_{game2,lose}\right), ... \right\}\right)$. We can generate a new task embedding $\hat{z}_{new} \in Z$:  

@@

Alternatively, we could map from language to a meta-mapping embedding, rather than inducing it from examples of the meta-mapping. This corresponds to the human ability to change behavior in response to instructions. The key feature of our architecture -- the fact that tasks, data, and language are all embedded in a shared space -- allows for substantial flexibility within a unified framework. Furthermore, our approach is parsimonious. Because it uses the same meta-learner for both basic tasks and meta-mappings, this increased flexibility does not require any added parameters.\footnote{At least in principle, in practice of course increasing network size might be more beneficial for EML architectures performing meta-mappings as well as basic tasks than for those performing only basic tasks.}

@@

\begin{figure}
\centering
\begin{subfigure}{0.5\textwidth}
\includegraphics[width=\textwidth]{figures/poly/meta_results.png}
\caption{The polynomials domain, section \ref{sec_poly}.}
\label{poly_meta_map_results_examples}
\end{subfigure}%
\begin{subfigure}{0.5\textwidth}
\includegraphics[width=\textwidth]{figures/meta_mapping.png}
\caption{The cards domain, section \ref{sec_cards}.}
\label{cards_meta_map_results_examples}
\end{subfigure}%
\caption{The EML architecture performs well at meta-mappings. (\subref{poly_meta_map_results_examples}) The system generalizes to apply learned meta-mappings to new polynomials, and to apply unlearned meta-mappings. The plots show the loss produced when evaluating the mapped embedding on the target task. For example, if the initial polynomial is $p(x) = x + 1$, and the meta-task is ``square,'' the loss would be evaluated by applying this meta-mapping to the embedding of $p(x)$ and evaluating how well the function specified by the mapped embedding regresses $p(x)^2 = x^2 + 2x + 1$. The solid line indicates optimal performance; the dashed line is chance (untrained).
(\subref{cards_meta_map_results_examples}) The system generalizes to meta-mapping new tasks in the cards domain. The system is trained to do the meta-mappings shown here on a subset of its basic tasks, and is able to generalize these mappings to novel tasks it has not been trained on. For example, for the ``losers'' mapping, the sytem is trained to map games to their losers variants. When given a novel game, it is able to apply this mapping to guess how to play the losers variation. This plot shows the reward produced by taking the mapped embedding and playing the targeted game. The gray dashed line indicates random performance, while the colored dashed lines indicate performance if the system did not alter its behavior in response to the meta-mapping. The system generally exceeds these baselines, although the switch-suits baseline is more difficult with the targeted holdout. 
Error-bars are bootstrap 95\%-CIs.} 
\label{meta_map_results}
\end{figure}

%% Cut example if necessary
As a proof of concept, we first evaluated the system on the task of learning polynomials of degree $\leq 2$ in 4 variables (i.e. the task was to regress functions of the form $p: \mathbb{R}^4 \rightarrow \mathbb{R}$ where $p \in \mathcal{P}_2 \left(\mathbb{R}\right)$, though the model was given no prior inductive bias toward polynomial forms). For example, if $p(w,x,y,z) = x$, the model might see examples like $(-1,1,1,1; 1)$ and $(0.7, 2.1, 1.3, -4; 2.1)$, and be evaluated on its output for points like $(-1, -1.3, 0.5, 0.3)$. This yields an infinite family of base-level tasks (the vector space of all such polynomials), as well as many families of meta-mappings over tasks (for example, multiplying polynomials by a constant, squaring them, or permuting their input variables). This allows us to not only examine the ability of the system to learn to learn polynomials from data, but also to adapt its learned representations in accordance with these meta-tasks. Details of the architecture and training can be found in appendix \ref{app_detailed_methods}.\par
\vspace{-0.7em}
\paragraph{Basic meta-learning:} First, we show that the system is able to achieve the basic goal of learning a held-out polynomial from a few data points in Fig. \ref{poly_basic_results} (with good sample-efficiency, see supp. Fig. \ref{supp_poly_sweep_results}). \par 
\vspace{-0.7em}
\paragraph{Meta-mapping (task $\rightarrow$ task):} Furthermore, the system is able to perform meta-mappings over polynomials in order to flexibly reconfigure its behavior, as shown in Fig. \ref{poly_meta_map_results}. We train the system to perform a variety of mappings, for example switch the first two inputs of the polynomial, add 3 to the polynomial, or square the polynomial. We then test its ability to generalize to held-out mappings, for example a permutation of the polynomial inputs which it has not encountered before, or an additive shift in the polynomial which it has not been trained on. The system is both able to apply its learned meta-mappings to held-out tasks, and to apply meta-mappings it has not been trained on, simply by seeing examples of the mapping. \par

@@

\section{Extensions}
%% TODO: language

\begin{figure}
\centering
\begin{subfigure}{0.5\textwidth}
\includegraphics[width=\textwidth]{figures/poly/language_meta_results.png}
\caption{From language input.}
\label{poly_meta_map_results_language}
\end{subfigure}%
\begin{subfigure}{0.5\textwidth}
\includegraphics[width=\textwidth]{figures/language_meta_mapping.png}
\caption{From language input.}
\label{cards_meta_map_results_language}
\end{subfigure}
\caption{The EML system can perform meta-mappings from language cues rather than meta-mapping examples, and can even generalize these. Compare to Fig. \ref{meta_map_results}, which shows the same results when using examples instead of language. (\subref{poly_meta_map_results_language}) In the polynomials domain, language cues still lead to good performance, even on held-out tasks or held-out meta-mappings, although examples perform slightly better (Fig. \ref{poly_meta_map_results_examples}). (\subref{cards_meta_map_results_language}) Similarly, in the cards domain, language cues perform well. Error-bars are bootstrap 95\%-CIs.}
\label{language_meta_map_results}
\end{figure}
\paragraph{Language:} Language is a fundamental piece of human flexibility, and we would be remiss to ignore it. In many of the intuitive illustrations we've given, the examples of the meta-mapping are implicit in your prior knowledge about the world. For example, ``try to lose at go'' does not give explicit examples of the ``lose'' meta-mapping, but rather relies on prior knowledge of what losing means. This is a much more efficient way to achieve zero-shot performance on a new losing task. In order to replicate this, we trained the EML system with both meta-mappings based on examples, and meta-mappings based on language. In the language-based meta-mappings, a language input identifying the meta-mapping (but not the basic task to apply it to) is provided to the network. This language input is encoded by a language encoder, and then provided as input to $\mathcal{H}$ instead of the output of $\mathcal{M}$. The meta-mapping then proceeds as normal --- $\mathcal{H}$ parameterizes $F$, which is used to transform the embedding of the input task to produce an embedding for the target task. \par
This language-cued meta-mapping approach also yields good performance (Fig. \ref{language_meta_map_results}). However, examples of the meta-mapping are slightly better, especially for meta-mappings not seen during training, presumably because examples provide a richer description. In Appendix \ref{app_why_meta_mapping} we show that using language to specify a meta-mapping performs better than using language to directly specify the target task, presumably by leveraging the richer task representation of the task embedding. \par

@@

@@

A very simple solution to these problems is naturally suggested by our architecture. Specifically, task embeddings can be cached so that they don't have to be regenerated at each inference step. This also allows optimization of these embeddings without altering the other parameters in the architecture, thus allowing fine-tuning on a task without seeing more examples, and without interfering with performance on any other task \citep[cf.][]{Rumelhart1993, Lampinen2018a}. That is, we can see the meta-learning step as a ``warm start'' for an optimization procedure over embeddings that are cached in memory \citep[cf.][]{Kumaran2016}.\footnote{Note that the weights generated by the hyper network could be cached instead of the embeddings, which would be much more memory intensive but would allow even better tuning and would not require passing through the hyper network at inference time.} While this is not a traditional continual learning perspective, we think it provides an interesting perspective on the issue.\par 
To test this idea, we pre-trained the system on 100 polynomial tasks, and then introduced 100 new tasks. We trained on these new tasks by starting from the meta-network's ``guess'' at the correct task embedding, and then optimizing this embedding without altering the other parameters. The results are shown in Fig. \ref{poly_continual_results}. The meta-network embeddings offer good immediate performance, and substantially accelerate the optimization process, compared to a randomly-initialized embedding (see supp. Fig. \ref{supp_poly_continual_results} for a more direct comparison). Furthermore, this ability to learn is due to training, not simply the expressiveness of the architecture, as is shown by attempting the same with an untrained network. \par 
